---
title: "Primary Care Model"
author: "Ed Young"
date:  '`r format(Sys.Date(), "%B %d, %Y")`'
output:
     html_document:
          theme: null
          highlight: null
          css: styles.css
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE, cache=FALSE, warning=FALSE, message=FALSE)
```



I am setting up a sample model using average demographics for New York State. I'll focus on gathering 
 - demographic data, 
 - estimating visit frequencies, and 
 - making initial predictions for resource requirements.

# Demographic Data
To build a representative model, I will need the following:

## Age distribution
based on the total population of New York State, which is approximately 19.6 million:

### Adults aged 18-44: Around 43.9%
### Adults aged 45-64: Around 34.5%
### Seniors aged 65 and over: Around 21.7%

## Gender distribution

## Common health issues or chronic conditions

# Estimate Visit Frequencies
Using the National Health Statistics Reports"Characteristics of Office-based Physician Visits by Age, 2019"  (https://www.cdc.gov/nchs/data/nhsr/nhsr184.pdf)  the estimate for the number of annual visits per demographic group are:

## Adults age 18 - 44: 1.88 visits per year
## Adults (45-64): 3.75 visits per year
## Seniors (65+): 7.09 visits per year

# Initial Predictions for Resource Requirements

```{r model setup}

# Sample demographic data based on New York State averages
age_group <- c("Age 18-44", "Age 45-64", "Age over 65")
percentage <- c(43.9, 34.5, 21.7)  # Hypothetical distribution
visits_per_year <- c(1.88, 3.75, 7.09)  # Hypothetical visit frequencies

# Create a data frame
df <- data.frame(Age_Group = age_group, Percentage = percentage, Visits_Per_Year = visits_per_year)

# Calculate the number of patients in each group
total_patients <- 10000
df$Number_of_Patients <- (df$Percentage / 100) * total_patients

# Calculate the total number of visits per year for each group
df$Total_Visits_Per_Year <- df$Number_of_Patients * df$Visits_Per_Year

# Display the data frame
df %>% 
  select(Age_Group, Percentage, Number_of_Patients, Visits_Per_Year, Total_Visits_Per_Year) %>% 
  kable

```
